---
title: "First graded assignment - Data Visualization"
format: html
author: Muriel Taks - mt3756
execute:
  echo: false
---

# How Couples Meet and Stay Together

Charts, interesting variables and suggestions regarding data visualization for Media Outlet, with data from "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

## Dating Trends over Time

```{r, warning=FALSE, message=FALSE, results='hide'}

HCMST_couples <- readRDS("C:/Users/Muriel/Downloads/HCMST_couples.rds")

library(ggplot2)
library(dplyr)
install.packages("plotly", repos="http://cran.r-project.org")
install.packages ("ggiraph", repos="http://cran.r-project.org")
install.packages("DT", repos="http://cran.r-project.org")
```

```{r, warning=FALSE, message=FALSE, results='hide'}

HCMST_couples$meeting_typerec <- ifelse(HCMST_couples$meeting_type %in% c("Online Gaming", "Met Online", "Internet Site", "Internet Chat", "Internet", "Internet Dating or Phone App", " Internet Social Network"), 1, 0)
freq_meeting_typerec <- table(HCMST_couples$meeting_typerec)

HCMST_couples$meeting_type2 <- ifelse(HCMST_couples$meeting_typerec == 1, "Online", "Offline")

meet_by_year <- HCMST_couples %>%
  group_by(Q21A_Year) %>%
  summarise(
    Online_Count = sum(meeting_typerec == 1),
    Offline_Count = sum(meeting_typerec == 0)
  )

meet_by_year <- na.omit(meet_by_year)

meet_by_year_df = as.data.frame(meet_by_year)
meet_by_year_df <- meet_by_year_df %>%
  filter(Q21A_Year != "Refused")

filtered_meet_by_year_df <- meet_by_year_df %>%
  filter(Q21A_Year %in% 1980:2017)
```

The first analysis regarding how couples meet throughout the years will be focused on two different ways of meeting a partner: Online and Offline. "Online" refers to those encounters that take place in vritual platforms such as Apps, Internet, Online Gaming and more. On the other hand, "Offline" refers to physical places or social frames such as Bars, Restaurants, School, Work and more, where potential couples interact. The time frame for this analysis was set between the years 1980 and 2017. The logic behind this decision is that, in order to see how the Online methods have grown, it would not be pertinent to include years like 1945, 1950 or 1967, where these methods did not exist yet.

```{r, warning=FALSE}

lines2 <- ggplot(filtered_meet_by_year_df, aes(x = Q21A_Year)) +
  geom_point(aes(y = Online_Count), color = "dodgerblue4") +
  geom_point(aes(y = Offline_Count), color = "maroon4") +
  geom_line(aes(y = Online_Count, group = 1), color = "dodgerblue3", size = 0.8) +
  geom_line(aes(y = Offline_Count, group = 1), color = "maroon3", size = 0.8) +
  labs(x = NULL, y = NULL, title = "Number of Online and Offline Meetings Across the Years (1980-2017)") +
  theme_minimal() +
  coord_cartesian(ylim = c(0, 100)) +
  scale_x_discrete(breaks = filtered_meet_by_year_df$Q21A_Year[seq(1, length(filtered_meet_by_year_df$Q21A_Year), by = 5)]) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))


lines3 <- lines2 + annotate("text", x = 5, y = 50, label = 
                              "Offline Meetings", 
                            fontface = "bold",
                            color = "maroon4", size = 3.5)

final_lines <- lines3 + annotate("text", x = 7, y = 10, label = 
                                   "Online Meetings", 
                                 fontface = "bold",
                                 color = "dodgerblue4", size = 3.5)

print (final_lines)
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

In this graph, I made the decision of including the label of each line ("Offline Meetings" and "Online Meetings") within the chart itself. This way, the reader will be able to go through all the information within the same visual space, without the need of going back and forth. Moreover, the labels of the years are displayed only every 5 years, so that the graph does not contain unnecessary information (you can understand the trend without the need of reading all the labels for the years from 1980 to 2017). Lastly, the names of both axis, which would have appeared as "Years" and "Frequencies" were erased, considering they are implicitly stated ("Frequencies" in the title of the plot, and "Years" only by reading the values of the x axis. All of these decisions were based on the principle of including only the indispensable information in the graph, so that to make the reading easier and more intuitive. Graphs too densely charged with information usually lead to confusion and lack of clarity.

```{r, warning=FALSE, message=FALSE, results='hide'}
#Analysis by month

meet_by_month <- HCMST_couples %>%
  group_by(Q21A_Month) %>%
  summarise(
    Online_Count = sum(meeting_typerec == 1),
    Offline_Count = sum(meeting_typerec == 0)
  )

#dropping NA

meet_by_month

meet_by_month <- na.omit(meet_by_month)

print(meet_by_month, n = 20)

#Conversion to dataframe

meet_by_month_df = as.data.frame(meet_by_month)
meet_by_month_df <- meet_by_month_df %>%
  filter(Q21A_Month != "Refused")

```

For the second visualization, the analysis was focused on how this same phenomenon evolves across the months of the year. The purpose was to highlight specifically the "Offline" kind of meetings, considering a significant pattern can be observed in the meetings of this type that take place in the month of September:

```{r, warning=FALSE}
meet_by_month_df <- transform(meet_by_month_df,
                              Offline_Count_Percentage = Offline_Count / sum(Offline_Count) * 100)


ggplot(meet_by_month_df, aes(x = Q21A_Month, y = Offline_Count_Percentage)) +
  geom_bar(stat = "identity", fill = "maroon4") +
  labs(x = NULL, y = "%", title = "How many people met Offline, by Month?") +
  theme_minimal() +
  ylim(0, 18) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.y = element_text(margin = margin(r = 10)),
         axis.text.x = element_text(size = 7))
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

As observed, this other chart reflects the development of only one category of the type of meeting, and it represents it throughout the years. As above mentioned, in this graph the label of the x axis was also erased (there is no need to clarify that those are months names), while the label of the y axis was displayed, but only clarifying that the frequencies are represented as percentages. In this case, the title was presented in the format of a question, to make the information displayed more appealing and straightforward. The high frequency of Offline meetings from June to September (with certain oscillations) might be explained by the nice weather during that time of the year in most of the cities in the United States, which might lead people to go out more and participate in outdoor activities, rather than using virtual platforms to meet others.

If I had to include only one of these visualizations, I would incline toward the first one. This is because this graph reflects a tendency throughout the years, allowing the reader to evidence a historical trend that might be of interest. The second graph, however, serves more like a "zoom in" in one of the categories, and it analyzes in a different time unit (which might also be interested for a deeper analysis, but not as a priority in the article).

## Age is just a Number

```{r, warning=FALSE, message=FALSE, results='hide'}
#Creating new variable that compares both ages (respondent's and partner's)

HCMST_couples$agecomparison <- ifelse(HCMST_couples$Q9 > HCMST_couples$ppage, 1, 0)
HCMST_couples$agecomparison <- ifelse(HCMST_couples$agecomparison == 1, "Partner is older", "Partner is younger or same age")

filtered_data <- subset(HCMST_couples, !(ppgender %in% c("Not asked", "refused")))

percentage_data <- with(filtered_data, table(agecomparison, ppgender) / sum(table(agecomparison, ppgender)) * 100)

percentage_df <- as.data.frame(percentage_data)
colnames(percentage_df) <- c("agecomparison", "ppgender", "percentage")
```

```{r}
graph_gender <- ggplot(subset(percentage_df, !(ppgender %in% c("Not asked", "refused"))), 
       aes(x = agecomparison, y = percentage, fill = ppgender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Partner's Age Comparison", y = "Percentage", fill = NULL) +
  ggtitle("Do women like their partners to be older than them?") +
  theme_minimal() +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(),
        axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  geom_text(aes(label = paste0(round(percentage), "%")),  
            position = position_dodge(width = 0.9), 
            vjust = -0.5, 
            size = 3,
            fontface = "bold")

graph_gender + annotate("text", x = 0.73, y = 17, label = 
                   "Only 12% of the cases 
where the partner
is older than the 
respondent are men", 
                 fontface = "bold",
                 color = "black", size = 2.5)
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

This new graph shows the relationship between a couples' ages (both the respondent's of the survey and his/her partner's) and the gender of the respondent. It is worth highlighting that, as the annotation suggests, whithin those cases where the partner is older than the respondent, only 12% of the respondents are men. Whereas, from this same group of people (partner is older than respondent), 32% identify themselves as women. The inverse relationship is observed in the second category (partner is younger or same age). A provocative title (presented as a question) was included, anticipating the result evidenced in the graph.

## Politics and Dating

```{r, warning=FALSE, message=FALSE, results='hide'}

#Creating new variable that compares respondent's and partner's political affiliation

levels(HCMST_couples$partyid7)
levels(HCMST_couples$w6_q12)

HCMST_couples$partyid7 <- as.character(HCMST_couples$partyid7)
HCMST_couples$w6_q12 <- as.character(HCMST_couples$w6_q12)

HCMST_couples$political <- ifelse(HCMST_couples$partyid7 == HCMST_couples$w6_q12, 1, 0)
HCMST_couples$political <- ifelse(HCMST_couples$political == 1, "Same political affiliation", "Different political affiliation")
table(HCMST_couples$political)

#New variable is called "political", let's see the relationship with quality of relationship

quality_political <- table(HCMST_couples$Q34, HCMST_couples$political)
quality_political_norm <- prop.table(quality_political, margin = 2) * 100
print(quality_political_norm)

normalized_data <- HCMST_couples %>%
  group_by(political, Q34) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)
```

```{r, warning=FALSE}
colors_Muriel <- c("NA" = "darkgray", "Very Poor" = "firebrick3", "Poor" = "firebrick1" , "Fair" = "lightgoldenrod1", "Good" = "darkolivegreen1", "Excellent" = "darkolivegreen3")

political1 <- ggplot(normalized_data, aes(x = political, y = percentage, fill = Q34)) +
  geom_bar(stat = "identity", position = "fill", width = 0.7) +
  labs(x = NULL, y = NULL, fill = NULL) +
  ggtitle("Quality of Relationship by Political Affiliation") +
  theme_minimal() +
  scale_fill_manual(values = colors_Muriel) + 
  theme(legend.position = "top", legend.text = element_text(angle = 0), axis.text.x = element_text(angle = 0, hjust = 0.5),  plot.title = element_text(face = "bold", hjust = 0.5))

print(political1)
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

This first visualization helps to depict the quality of the relationship, stated by the respondent of the survey, in relationship to the political affiliation of the couple. A new, recoded variable with two categories was built for the purpose of this analysis regarding coincidence/difference of political affilitation. These two new categories are; couples with different political affiliation and couples with same political affiliation. It can be observed that the quality of the relationship (that is represented by a scale that goes from "Very poor" to "Excellent") does not vary as much as one would imagine between those couples that differ in terms of political affiliation (I would have anticipated that within the couples who differed there would be higher levels of "Fair" and "Poor" relationships, for example).

A second visualization was created, representing the same variables but varying slightly (yet significantly in terms of interpretation) some aspects of the visualization.

```{r, warning=FALSE}
political2 <- ggplot(normalized_data, aes(x = percentage, y = political, fill = Q34)) +
  geom_bar(stat = "identity", position = "fill", width = 0.7) +
  geom_text(data = subset(normalized_data, Q34 %in% c("Excellent")), 
            aes(label = paste0(round(percentage), "%")), 
            position = position_fill(vjust = 0.8), 
            size = 3, 
            fontface = "bold") +
  labs(x = NULL, y = NULL, fill = NULL) +
  ggtitle("Quality of Relationship by Political Affiliation") +
  theme_minimal() +
  scale_fill_manual(values = colors_Muriel) + 
  theme(legend.position = "top", 
        legend.text = element_text(angle = 0), 
        axis.text.y = element_text(hjust = 0.5),  
        plot.title = element_text(face = "bold"))

print(political2)
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

The graph was changed to a horizontal orientation considering the proportions can be easier compared to one another this way. In this new kind of visualization, the differences between the categories of "Quality of Relationship" for each group can be clearly evidenced. Moreover, a label for the category "Excellent" was added to the charts, so that the most extreme category of the quality of the relationship could be observed easily. This way, the comparison between couples with same and different political affiliation can be done faster by the reader. For all of this, I would recommend including this second variation in the article.

## Potential topic:

## Who wanted to break up more?

**Differences between kinds of couples**

```{r, warning=FALSE, message=FALSE, results='hide'}

cross_tab2 <- table(HCMST_couples$w6_breakup_nonmar, HCMST_couples$Q5_2)
frequency_crosstab2 <- as.data.frame(cross_tab2) %>%
  rename(w6_breakup_nonmar = Var1, Q5_2 = Var2, Frequency = Freq)

frequency_crosstab2 <- frequency_crosstab2 %>%
  group_by(w6_breakup_nonmar) %>%
  mutate(Frequency = Frequency / sum(Frequency) * 100)
```

```{r}
ggplot(frequency_crosstab2 %>% filter(Q5_2 != "Refused" & w6_breakup_nonmar != "Refused"),
       aes(x = Q5_2, y = w6_breakup_nonmar, fill = Frequency)) +
  geom_tile() +
  theme_minimal() +
  scale_fill_gradient(low = "ghostwhite", high = "dodgerblue", name = "Percentage") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"), axis.text.x = element_text(size = 6.5)) +
  labs(x = NULL, y = NULL, title = "Decision of Breakup by Type of Couple")
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

This graph intends to visualize the influence of the type of couple (same-sex or opposite-sex) on the decision of the break up. As the heat map shows, within the same-sex couple there is a higher tendency for the respondent to affirm that he/she wanted to break up more than his/her partner. This might be an interesting topic for the readers, considering is related to the break up style, as well as sexual identity, two subjects that might catch the eye of the audience.

### Another Potential topic:

## Who has sex more frequently? Same-sex or opposite-sex couples?

```{r, warning=FALSE, message=FALSE, results='hide'}

#Analyzing sex frequency by types of couples

frequency_table <- table(HCMST_couples$w6_sex_frequency, HCMST_couples$Q5)

frequency_df <- as.data.frame(frequency_table)
names(frequency_df) <- c("w6_sex_frequency", "Q5", "Frequency")

frequency_df <- frequency_df %>%
  group_by(Q5) %>%
  mutate(Relative_Frequency = Frequency / sum(Frequency) * 100)

frequency_df <- frequency_df %>%
  filter(Q5 != "Refused")
```

```{r, warning=FALSE}
ggplot(frequency_df, aes(x = Q5, y = Relative_Frequency, fill = w6_sex_frequency)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  scale_fill_brewer(palette = "YlOrRd") +
  labs(x = NULL, y = NULL, fill = "Sex Frequency") +
  ggtitle("Sex frequency by Type of Couple") + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5)) +
  geom_text(aes(label = ifelse(w6_sex_frequency %in% c("3 to 6 times a week", "Once or twice a week", "2 to 3 times a month"), 
                               paste0(round(Relative_Frequency), "%"), "")),
            position = position_stack(vjust = 0.5), 
            size = 3, fontface = "bold", color = "black")
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

This graph analyzes a different phenomenon that relates to the kind of couple; frequency of sex. Labels on some of the sex frequency variable's categories were created to make the comparison easier. As observed, the higher frequency (Once a day or more) shows a slightly higher value in the opposite-sex couples. One of the lowest categories of sex frequency (2 to 3 times a month) is also more reduced in this group of people, in comparison to same-sex couples.

If I had to suggest which of these two graphs to present, I would go for this second one. Although the break up decision might be an interesting piece of information for certain audiences, I believe readers in general might be more attracted to data related to sexual activity. Also, this could contribute to weaken some strong prejudices toward same-sex couples and their "intense" sexual habits, widely spread in our society.

## Interactive Plots

```{r, warning=FALSE, message=FALSE, results='hide'}

library(plotly)
```

```{r}
#Making the final_lines (Offline and Online through the years) interactive

ggplotly(final_lines)
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

Returning to the first analysis of this study, it might be helpful to introduce an interactive mode of the "Offline"-"Online" trends throughout the years. First, this tool allows the reader to better understand the trend in each of the year that he/she is interested in. For aesthetic reasons, as it was previously explained, the x axis only displays the years with a 5-unit break. However, someone might be interested in seeing how the ways of meeting partners were in a specific year. This is possible in an interactive chart. Secondly, the interactive visualization allows to read the accurate count for each of the meeting modes, which might not be an easy task only by guiding the reading by the y axis information. So, for example, if the reader wants to know how many people who met in 1999 did so through online methods, he/she will be able to through this visualization.

```{r}
#Making the political2 graph (quality of relationship by political affiliation) interactive

library(ggiraph)

political_interactive <- ggplot(normalized_data, aes(x = percentage, y = political, fill = Q34, 
                                                     tooltip = paste("Percentage:", round(percentage, digits = 0)))) +
  geom_bar_interactive(stat = "identity", position = "fill", width = 0.7) +
  labs(x = NULL, y = NULL, fill = NULL) +
  ggtitle("Quality of Relationship by Political Affiliation") +
  theme_minimal() +
  scale_fill_manual(values = colors_Muriel) + 
  theme(legend.position = "top", 
        legend.text = element_text(angle = 0), 
        axis.text.y = element_text(hjust = 0.5),  
        plot.title = element_text(face = "bold"))


ggiraph(code = print(political_interactive))

```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

This second interactive visualization allows the reader to know the exact percentage that corresponds to each category of the variable "Quality of the Relationship", differentiated by each of the groups of political affiliation. This way, an easier and more intuitive comparison can be made between the two types of couples (those with same political affiliation and those with different), if the quality of their bond wants to be analyzed.

## Interactive Data Table

```{r, warning=FALSE, message=FALSE, results='hide'}

#Creating new dataframe

interactive_df <- HCMST_couples[c("ppgender", "Q34", "meeting_type", "political", "Q5_2", "w6_sex_frequency")]
head(interactive_df)

#Renaming variables 

interactive_df <- interactive_df %>%
  rename(
"Gender of respondent" = ppgender,
"Relationship Quality" = Q34,
"Political Affiliation of Couple" = political,
"Place of couple's meeting" = meeting_type,
"Type of couple" = Q5_2,
"Sex frequency" = w6_sex_frequency)
```

```{r, warning=FALSE}

library(DT)

interactive_df %>% 
  datatable(caption = "This table recaps some of the variables previously analyzed. Let's be honest, we all love to know about other couples' dynamics",
            filter = "top",
            options = list(
              pageLength = 20,
              dom = "Bftrip",
              buttons = c("copy", "csv", "excel", "pdf")
            ),
            extensions = "Buttons"
  )
```

Source: "How Couples Meet and Stay Together" (Rosenfeld, Reuben & Falcon, 2018).

Lastly, this data table was created for the reader to be able to examine and browse through the data as he/she wishes. As can be read in the header of the table, we have to acknowledge that everyone can (and in my opinion, is) curious about other people's relationship.

To facilitate this process, filters in each columns were added, so that the audience can directly type a specific value (for example, "Excellent" in the variable "Relationship Quality") and observe how the other variables are modified based on this filtering. Plus, "Copy", "CSV", "Excel" and "PDF" buttons were also added so that the reader is free to download the data in the preferred format. Of course, this should be discussed with the Media Outlet, however, from my perspective, it is always a good idea to share the datasets behind the published information and allow others to manipulate, analyze or simply observe the data.
